---
title: "tree outputs"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(svglite)
library(ggtree)
library(ape)

```

Instructions for quest 

1. extract seqs of interest from rep_seqs file (qiime output)
module load seqkit (version is 0.6.0)
cd /projects/p31629/zhen_check
seqkit grep --pattern-file ./intermediate/acc_unique.txt ./unzipped_qza/rep_seqs_all.fasta -o ./intermediate/acc_only.fa
seqkit grep --pattern-file ./intermediate/azo_unique.txt ./unzipped_qza/rep_seqs_all.fasta -o ./intermediate/azo_only.fa

2. blast and save seqs
dedup seqs from blast results
cd ./intermediate
seqkit rmdup -n -i acc_blast.fa -o acc_blast_dedup.fa
(skip for azo)

3. prep for msa
add qiime seqs to dedup file
add outgroup
cat acc_blast_dedup.fa acc_only.fa root.fa /projects/b1052/mckenna/resources/midas_5.3/acc_trim_clean_dedup.fa > acc_for_align.fa
seqkit rmdup -n -i acc_for_align.fa -o acc_for_align_dedup.fa

cat azo_only.fa root.fa /projects/b1052/mckenna/resources/midas_5.3/azo_trim_clean_dedup.fa > azo_for_align.fa
seqkit rmdup -n -i azo_for_align.fa -o azo_for_align_dedup.fa


4. msa
module load mafft (7.407)
mafft acc_for_align_dedup.fa > acc_align.msa
mafft azo_for_align_dedup.fa > azo_align.msa


module load fasttree (2.1.10)
FastTree acc_align.msa > acc_tree.tre
FastTree azo_align.msa > azo_tree.tre




```{r fig.height = 10}
acc_tree <- read.tree("./intermediate/acc_tree.tre")
acc_tree_root <- root(acc_tree, outgroup = "NR_114042.1", edgelabel = TRUE, resolve.root = TRUE)

ggtree(acc_tree_root) +
  geom_nodepoint(aes(subset=acc_tree_root$node.label>=0.9), size = 2) +
  geom_nodepoint(aes(subset=acc_tree_root$node.label<0.9 & acc_tree_root$node.label>=0.5), size = 2, color = "blue", shape = 1) +
  geom_tree() +
  geom_tiplab() +
  theme_tree() + 
  geom_treescale()


ggsave("./figures/acc_test.svg", width = 60, height = 10, units = "in", limitsize = FALSE)
```

```{r fig.height = 10}
azo_tree <- read.tree("./intermediate/azo_tree.tre")
azo_tree_root <- root(azo_tree, outgroup = "NR_114042.1", edgelabel = TRUE, resolve.root = TRUE)

# ggtree(azo_tree_root) +
#   geom_nodepoint(aes(subset=azo_tree_root$node.label>=0.9), size = 2) +
#   geom_nodepoint(aes(subset=azo_tree_root$node.label<0.9 & azo_tree_root$node.label>=0.5), shape=10, size = 2) +
#   geom_tree() +
#   geom_tiplab() +
#   theme_tree()


midas_taxonomy <- read_delim("~/b1052/mckenna/resources/midas_5.3/midas_5.3.txt", 
                             col_names = c("kingdom", "phylum", "class", 
                                           "order", "family", "genus", "species")) %>%
    separate(col = kingdom, into = c("id", "kingdom"), sep = "\\s") %>%
    mutate_at(vars(kingdom:species), ~ str_replace(., ";", "")) %>%
    mutate_at(vars(kingdom:species), ~gsub(x = ., "[A-z]__", ""))

azo_labs <- as.data.frame(list(azo_tree$tip.label), col.names = c("id")) %>%
  left_join(midas_taxonomy, by = "id") %>%
  mutate(id_new = case_when(grepl("midas_s", species) ~ id,
                            is.na(species) ~ id,
                            .default = paste0(id, " (", species, ")")))

azo_tree_root$tip.label <- azo_labs$id_new


ggtree(azo_tree_root) +
  geom_nodepoint(aes(subset=azo_tree_root$node.label>=0.9), size = 2) +
  geom_nodepoint(aes(subset=azo_tree_root$node.label<0.9 & azo_tree_root$node.label>=0.5), size = 2, color = "blue", shape = 1) +
  geom_tree() +
  geom_tiplab() +
  theme_tree()+ 
  geom_treescale()

ggsave("./figures/azo_test.svg", width = 60, height = 20, units = "in", limitsize = FALSE)
```


