---
title: "ASV analysis for PAO and GAO"
output:
  html_document:
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


```{r packages}

library(tidyverse)

source("./scripts/plotting.R")
```


Data read in and parsing
```{r}

table_rel_long <- readRDS("./intermediate/table_rel_long")

midas_taxonomy <- read_delim("~/b1052/mckenna/resources/midas_5.3/midas_5.3.txt", 
                             col_names = c("kingdom", "phylum", "class", 
                                           "order", "family", "genus", "species")) %>%
    separate(col = kingdom, into = c("id", "kingdom"), sep = "\\s") %>%
    mutate_at(vars(kingdom:species), ~ str_replace(., ";", "")) %>%
    mutate_at(vars(kingdom:species), ~gsub(x = ., "[A-z]__", ""))


```




# Competibacter denitrificans?
No matches with Competibacter denitrificans sequences
```{r message=FALSE, warning=FALSE}
blast_bc_comp <- read_delim("./blast_out/competibacter_blast_bc.txt", 
                            col_names = c("query", "subject", "pident", "length", 
                            "mismatch", "gapopen", "qstart", "qend",
                            "sstart", "send", "evalue", "bitscore")) %>%
  filter(pident >= 97) %>%
  left_join(midas_taxonomy, by = c("subject" = "id"))

blast_ss_1_comp <- read_delim("./blast_out/competibacter_blast_ss_1.txt", 
                            col_names = c("query", "subject", "pident", "length", 
                            "mismatch", "gapopen", "qstart", "qend",
                            "sstart", "send", "evalue", "bitscore")) %>%
  filter(pident >= 97) %>%
  left_join(midas_taxonomy, by = c("subject" = "id"))

blast_ss_2_comp <- read_delim("./blast_out/competibacter_blast_ss_2.txt", 
                            col_names = c("query", "subject", "pident", "length", 
                            "mismatch", "gapopen", "qstart", "qend",
                            "sstart", "send", "evalue", "bitscore")) %>%
  filter(pident >= 97) %>%
  left_join(midas_taxonomy, by = c("subject" = "id"))

blast_comp_all <- bind_rows(blast_bc_comp, blast_ss_1_comp, blast_ss_2_comp) %>%
  distinct(across(everything()))

rm(blast_bc_comp, blast_ss_1_comp, blast_ss_2_comp)

```


# Dechloromonas PAO?
A few matches to Dechloromonas phosphoritropha. I also looked for Dechloromonas denitrificans, agitata, and hortensis but no significant hits. That doesn't necessarily mean there weren't denitrifying Dechloromonas though, just that we couldn't determine based on 16S.
```{r message=FALSE, warning=FALSE}
blast_bc_dechloro <- read_delim("./blast_out/dechloro_blast_bc.txt", 
                            col_names = c("query", "subject", "pident", "length", 
                            "mismatch", "gapopen", "qstart", "qend",
                            "sstart", "send", "evalue", "bitscore")) %>%
  filter(pident >= 97) %>%
  left_join(midas_taxonomy, by = c("subject" = "id"))

blast_ss_1_dechloro <- read_delim("./blast_out/dechloro_blast_ss_1.txt", 
                            col_names = c("query", "subject", "pident", "length", 
                            "mismatch", "gapopen", "qstart", "qend",
                            "sstart", "send", "evalue", "bitscore")) %>%
  filter(pident >= 97) %>%
  left_join(midas_taxonomy, by = c("subject" = "id"))

blast_ss_2_dechloro <- read_delim("./blast_out/dechloro_blast_ss_2.txt", 
                            col_names = c("query", "subject", "pident", "length", 
                            "mismatch", "gapopen", "qstart", "qend",
                            "sstart", "send", "evalue", "bitscore")) %>%
  filter(pident >= 97) %>%
  left_join(midas_taxonomy, by = c("subject" = "id"))


blast_dechloro_all <- bind_rows(blast_bc_dechloro, blast_ss_1_dechloro, blast_ss_2_dechloro) %>%
  distinct(across(everything()))

blast_dechloro_pao <- blast_dechloro_all %>%
  filter(species == "Azonexus_phosphoritropha")

rm(blast_bc_dechloro, blast_ss_1_dechloro, blast_ss_2_dechloro)


saveRDS(blast_dechloro_pao, file = "./intermediate/blast_dechloro_pao.RDS")

```


```{r}


table_rel_long %>% 
  filter(Genus == "Azonexus") %>%
  mutate(dechloro_pao = ifelse(seq %in% blast_dechloro_pao$query, "PAO", "Non-PAO")) %>%
  group_by(day, location, dechloro_pao) %>%
  summarise(sum_ab = sum(rel_ab)) %>%
  ggplot(., aes(x = day, y = sum_ab, color = location, shape = location)) +
  facet_wrap(~dechloro_pao, nrow = 2, scales = "free") +
  phases_rect +  
  geom_point() +
  geom_line() +
  scale_color_manual(values = location_cols, name = "Location") + 
  scale_x_continuous(breaks = seq(0, 300, by = 50)) + 
  scale_shape_manual(values = c(15, 16), name = "Location") +
  labs(x = "Day of Operation", y = "Relative abundance [%]") + 
  theme_classic() +
  theme(legend.position = "top")

ggsave("./figures/azonexus_pao_panels.png", width = 4, height = 5, units = "in", dpi = 320)


azo_only <- table_rel_long %>%
  filter(Genus == "Azonexus")

most_ab <- azo_only %>%
  group_by(seq) %>%
  summarise(median = median(rel_ab)) %>%
  filter(median > 0 )

azo_only_unique <- most_ab %>%
  select(seq) %>%
  ungroup() %>%
  distinct(seq)

azo_only %>%
  filter(seq %in% most_ab$seq) %>%
  group_by(day, location, seq) %>%
  summarise(sum_ab = sum(rel_ab)) %>%
  ggplot(., aes(x = day, y = sum_ab, color = seq, shape = seq)) +
  facet_wrap(~location, nrow = 2) +
  phases_lines + 
  geom_point(alpha = 0.7, size = 2) +
  geom_line() + 
  scale_x_continuous(breaks = seq(0, 300, by = 30)) + 
  labs(x = "Day of Operation", y = "Relative abundance [%]", title = "Azonexus") + 
  theme_classic()


write_delim(azo_only_unique, "./intermediate/azo_unique.txt", col_names = FALSE)


```


# Competibacter ASV analysis

```{r fig.width = 10, fig.height = 6}

comp_only <- table_rel_long %>%
  filter(Genus == "Ca_Competibacter")

most_ab <- comp_only %>%
  group_by(seq) %>%
  summarise(median = median(rel_ab)) %>%
  filter(median > 0 )

comp_only %>%
  filter(seq %in% most_ab$seq) %>%
  group_by(day, location, seq) %>%
  summarise(sum_ab = sum(rel_ab)) %>%
  ggplot(., aes(x = day, y = sum_ab, color = seq, shape = seq)) +
  facet_wrap(~location, nrow = 2) +
  phases_lines + 
  geom_point(alpha = 0.7, size = 2) +
  geom_line() + 
  scale_x_continuous(breaks = seq(0, 300, by = 30)) + 
  labs(x = "Day of Operation", y = "Relative abundance [%]", title = "Competibacter ASVs") + 
  theme_classic()

comp_only_unique <- most_ab %>%
  select(seq) %>%
  ungroup() %>%
  distinct(seq)

write_delim(comp_only_unique, "./intermediate/comp_unique.txt", col_names = FALSE)

```


```{r fig.width = 10, fig.height = 6}
acc_only <- table_rel_long %>%
  filter(Genus == "Ca_Accumulibacter")

most_ab <- acc_only %>%
  group_by(seq) %>%
  summarise(median = median(rel_ab)) %>%
  filter(median > 0.2)

acc_only %>%
  filter(seq %in% most_ab$seq) %>%
  group_by(day, location, seq) %>%
  summarise(sum_ab = sum(rel_ab)) %>%
  ggplot(., aes(x = day, y = sum_ab, color = seq, shape = seq)) +
  facet_wrap(~location, nrow = 2) +
  phases_lines + 
  geom_point(alpha = 0.7, size = 1) +
  geom_line() + 
  scale_x_continuous(breaks = seq(0, 300, by = 30)) + 
  labs(x = "Day of Operation", y = "Relative abundance [%]", title = "Accumulibacter ASVs") + 
  theme_classic()

acc_only_unique <- most_ab %>%
  select(seq) %>%
  ungroup() %>%
  distinct(seq)

write_delim(acc_only_unique, "./intermediate/acc_unique.txt", col_names = FALSE)

```

